### 2008-04-22 opal/documents/grid2008/paper.tex
added a template with the correct formatting

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}

\begin{document}

% paper title
\title{On Supporting Molecular Visualization on Grid Resources using Visual
  Programming Environments}

% author information
\author{\authorblockN{Sriram Krishnan\authorrefmark{1}\authorrefmark{2},
    Guillaume Vareille\authorrefmark{3},
    Wesley Goodman\authorrefmark{1},
    Luca Clementi\authorrefmark{1},\\
    Jane Ren\authorrefmark{1},
    Sargis Dallakyan\authorrefmark{3},
    Wilfred Li\authorrefmark{1}\authorrefmark{2},
    Michel Sanner\authorrefmark{3} and
    Peter Arzberger\authorrefmark{1}}
  \authorblockA{\authorrefmark{1}National Biomedical Computation Resource\\
    CRBS, UC San Diego, 9500 Gilman Dr, La Jolla, CA 92093}
  \authorblockA{\authorrefmark{2}San Diego Supercomputer Center\\
    UC San Diego MC 0505, 9500 Gilman Dr, La Jolla, CA 92093}
  \authorblockA{\authorrefmark{3}The Molecular Graphics Lab\\
    The Scripps Research Institute, La Jolla, CA 92093
    }
  \authorblockA{sriram@sdsc.edu, vareille@scripps.edu, \{wgoodman,
      lclement, j2ren@ucsd.edu\},\\
      wilfred@sdsc.edu, sanner@scripps.edu, parzberger@ucsd.edu}
  }

% make the title area
\maketitle

\begin{abstract}

  Abstract goes here (Sriram)

\end{abstract}

\section{Intro on scientific workflows and grid computing}

- Scientific workflows: Michel

- Grid computing: Sriram

\section {Vision vs other workflow management systems (MS, except section
  2.c)}

- Design -- workflow representation and control: Michel

- API -- how to write Vision nodes, introspection, etc: Michel

- Comparison with other tools: Luca

  - Kepler, Ptolemy and MoML

  - Taverna, Scufl and Freeflou

\section {Application based services}

- Opal wrapper and GUI \cite{opal}: Sriram

- Opal and Opal-CSF4: Sriram

- Data services: Sriram

\section {Service based workflows}

- Vision based grid workflows (Michel, Guillaume, Sargis)

- Virtualization of Opal services (Sargis - something about how you
automatically create Vision nodes from the service list)

- Virtualization of Vision based workflows

  i) Use a local CLI opal client to access local or remote services (Sriram)

 ii) Automatic vision node creation through the opal client information
 service (Sriram, Guillaume - how the generated class is converted into a Vision node using introspection)

\section {Discussion - Workflow modeling and execution}

- How vision currently stores workflow (Michel)

- Fault tolerance (Michel, Sriram)

- Support for XPDL and other representations for workflows (Luca)

\section {Conclusions}

- Conclusions (Sriram)

%% References
\bibliographystyle{plain}
\bibliography{paper}

% that's all folks
\end{document}