### 2008-04-23 opal/documents/grid2008/paper.tex
added abstract

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}

\begin{document}

% paper title
\title{On Supporting Molecular Visualization on Grid Resources using Visual
  Programming Environments}

% author information
\author{\authorblockN{Sriram Krishnan\authorrefmark{1}\authorrefmark{2},
    Guillaume Vareille\authorrefmark{3},
    Wesley Goodman\authorrefmark{1},
    Luca Clementi\authorrefmark{1},\\
    Jane Ren\authorrefmark{1},
    Sargis Dallakyan\authorrefmark{3},
    Wilfred Li\authorrefmark{1}\authorrefmark{2},
    Michel Sanner\authorrefmark{3} and
    Peter Arzberger\authorrefmark{1}}
  \authorblockA{\authorrefmark{1}National Biomedical Computation Resource\\
    CRBS, UC San Diego, 9500 Gilman Dr, La Jolla, CA 92093}
  \authorblockA{\authorrefmark{2}San Diego Supercomputer Center\\
    UC San Diego MC 0505, 9500 Gilman Dr, La Jolla, CA 92093}
  \authorblockA{\authorrefmark{3}The Molecular Graphics Lab\\
    The Scripps Research Institute, La Jolla, CA 92093
    }
  \authorblockA{sriram@sdsc.edu, vareille@scripps.edu, \{wgoodman,
      lclement, j2ren@ucsd.edu\},\\
      wilfred@sdsc.edu, sanner@scripps.edu, parzberger@ucsd.edu}
  }

% make the title area
\maketitle

\begin{abstract}

  The visual programming paradigm provided by tools such as AVS [ref],
  OpenDX [ref], SciRun [ref], etc have been around for a couple decades.
  Recently, with the advent of the Web 2.0 paradigm and the concept of {\em
    mashups} [ref], there has been renewed interest in such types of
  environments. Over the past several years, we have developed one such
  tool called Vision [ref], which is a component-based, application-domain
  agnostic, cross-platform visual programming environment. Meanwhile,
  advances in distributed and Grid computing have opened up avenues for
  scientists to run large simulations on a scale that was impossible until
  recent times. In the past, visual programming tools were generally
  limited to running scientific experiments on user's individual
  workstations, and had limited capability to use distributed resources.
  With the proliferation of Web services, distributed resources are much
  more easy to access from any workstation that is connected to the
  Internet. In particular, we have developed the Opal toolkit, which
  provides a mechanism to wrap scientific applications running on Grid
  resources as Web services, thus providing the clients with a simple Web
  API to the back-end computational functionality. Opal encapsulates
  standard Grid security mechanisms, schedulers, and state management for
  jobs, and also provides a mechanism to describe command-line arguments
  that can be used to automatically generate user-interfaces [ref]. In this
  paper, we describe our current work on enabling access to scientific
  applications running on Grid resources via Opal from the Vision
  environment. We present brief overviews of both Vision and Opal, and
  discuss a scientific workflow that has been implemented with the above
  tools. We discuss our experiences, and plans for future work.

\end{abstract}

\section{Intro on scientific workflows and grid computing}

- Scientific workflows: Michel

- Grid computing: Sriram

\section {Vision vs other workflow management systems (MS, except section
  2.c)}

- Design -- workflow representation and control: Michel

- API -- how to write Vision nodes, introspection, etc: Michel

- Comparison with other tools: Luca

  - Kepler, Ptolemy and MoML

  - Taverna, Scufl and Freeflou

\section {Application based services}

- Opal wrapper and GUI \cite{opal}: Sriram

- Opal and Opal-CSF4: Sriram

- Data services: Sriram

\section {Service based workflows}

- Vision based grid workflows (Michel, Guillaume, Sargis)

- Virtualization of Opal services (Sargis - something about how you
automatically create Vision nodes from the service list)

- Virtualization of Vision based workflows

  i) Use a local CLI opal client to access local or remote services (Sriram)

 ii) Automatic vision node creation through the opal client information
 service (Sriram, Guillaume - how the generated class is converted into a Vision node using introspection)

\section {Discussion - Workflow modeling and execution}

- How vision currently stores workflow (Michel)

- Fault tolerance (Michel, Sriram)

- Support for XPDL and other representations for workflows (Luca)

\section {Conclusions}

- Conclusions (Sriram)

%% References
\bibliographystyle{plain}
\bibliography{paper}

% that's all folks
\end{document}